Embedding Intuitionistic into Classical Logic

### 2022-07-14 CSL/lipics-v2021-EIICL.tex
typo
added some comments

@@

Before proceeding towards the results let us quickly fix our notation and recapitulate the semantics for classical and more importantly intuitionistic logic.

@@

\begin{definition}
	A \emph{valuation} $v$ is a function that assigns to each propositional variable $A$ a \textbf{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulas:
	\begin{itemize}
		\item $v\not\models \bot$
		\item $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
		\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
		\item $v\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
		\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ or $v\models\psi$.
	\end{itemize}
	$v$ is a \textbf{model} for $\varphi$ if $v\models\varphi$. If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}. We denote the set of valid formulas with \emph{CPC}.
\end{definition}

@@

\begin{definition}
	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of nodes an a family of valuations $(v_w)_{w\in W}$ such that for $u\leq w$ we have $v_u(A)\leq v_u(A)$ for every propositional variable $A$, this is called the \emph{persistency condition}. The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, i.e. have
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
	\end{itemize}
	We say that $\varphi$ is satisfied at a node $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every node then $\mathcal K$ is a \textbf{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulas with IPC.
\end{definition}
There are many classical tautologies which are not intuitionistically valid
\begin{example}\label{LEMcounterexample}
	Consider the famous law of excluded middle $A\vee\neg A$. Recall that $\neg A := A\to \bot$. As a counter-model consider a Kripke structure with two nodes $u\leq w$ and $v_u(A) = 0, v_w(A) = 1$. Then $u\not\models A$ and $u\not\models \neg A$, i.e. $u\not\models A\vee\neg A$. 
\end{example}
In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable. This is insufficient for intuitionistic logic, e.g. $A\vee\neg A$ is not a intuitionistic tautology but its negation is not satisfiable either.

@@

We now give account of the more involved semantics of first-order logic

\begin{definition}
	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the domain of $\mathcal{M}$, and a function $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
		m&\text{if $b=a$}\\
		v(b)&\text{otherwise}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$.
		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
		\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$. We denote the set of valid formulas with CQC.
\end{definition}

@@

\subsection{Skolemization and Herbrandization}
An important step in the embedding will be the elimination of quantifiers via Herbrandization. In this process we introduce free variables and also additional function symbols to the signature. Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas. Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.

@@

\begin{theorem}For every formula $\varphi$
	\begin{itemize}
		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
		\item $\varphi$ and $\varphi^H$ are classically equivalid.
	\end{itemize}
\end{theorem}

which follows from Lemmata~\ref{ap1} and~\ref{ap2} in the appendix. We are now ready to present our embedding of intuitionistic into classical logic.

@@

\subsection*{Reducing the encoding}

>>>>

\usepackage{todonotes}

@@

In this section we fix notation and recall the semantics for classical and intuitionistic logic.

@@

\begin{definition}
A \emph{valuation} $v$ maps each propositional variable $A$ to a \emph{truth value} $v(A)\in\{0, 1\}$. We inductively define the model relation $\models$ between $v$ and formulas:
	\begin{itemize}
		\item $v\not\models \bot$
		\item $v\models A$ iff $v(A) = 1$ for each propositional variable $A$.
		\item $v\models \varphi\wedge\psi$ iff $v\models\varphi$ and $v\models\psi$.
		\item $v\models\varphi\vee\psi$ iff $v\models\varphi$ or $v\models\psi$.
		\item $v\models\varphi\to \psi$ iff $v\not\models\varphi$ or $v\models\psi$.
	\end{itemize}
A valuation $v$ is a \emph{model} for $\varphi$ if $v\models\varphi$. If every valuation is a model for $\varphi$ then we say $\varphi$ is \emph{valid}. We denote the set of valid formulas with \emph{CPC}.
\todo{What does CPC stand for? Introduce abbreviation somewhere like classical predicate ... (CPC).}
\end{definition}

@@

\begin{definition}
\todo{We later also use the notion `Kripke frame', we  might introduce it here.}
\todo{Note that I have replaced `nodes' by `worlds' here.}
A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds} and a family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_u(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
The model relation between the $v_u$ and formulas $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ if for all $w\geq u$ we have $w\not\models\varphi$ or $u\models\psi$.
	\end{itemize}
We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulas with IPC.
\todo{introduce abbreviation}
\end{definition}
There are many classical tautologies which are not intuitionistically valid:
\begin{example}\label{LEMcounterexample}
Consider the famous law of excluded middle $A\vee\neg A$. Recall that $\neg A := A\to \bot$. As a counter-model consider a Kripke structure with two nodes $u\leq w$ and $v_u(A) = 0, v_w(A) = 1$. Then $u\not\models A$ and $u\not\models \neg A$, i.e. $u\not\models A\vee\neg A$.
\end{example}
In classical logic a proven strategy for showing that a formula valid is to consider its negation and show that it is not satisfiable. This is insufficient for intuitionistic logic, e.g. $A\vee\neg A$ is not an intuitionistic tautology but its negation is not satisfiable either.

@@

We now recall the semantics of first-order logic:

\begin{definition}
Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary relation symbol $R$ a $n$-ary relation $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
		m&\text{if $b=a$}\\
		v(b)&\text{otherwise}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$.
		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$.
		\item $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$ where a is a free variable that does not occur in $\varphi$.
	\end{itemize}
	$\mathcal{M}$ satisfies $\varphi$ if $\mathcal M, v\models\varphi$ for every $v$. We denote the set of valid formulas with CQC.
\end{definition}

@@

\subsection{Skolemization and Herbrandization}

An important step in the embedding will be the elimination of quantifiers via Herbrandization. In this process we introduce free variables and also additional function symbols to the signature. Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas. Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.

@@

\begin{theorem}
\label{thm:herbrand-skolem}
For every formula $\varphi$
	\begin{itemize}
		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
		\item $\varphi$ and $\varphi^H$ are classically equivalid.
	\end{itemize}
\end{theorem}

Theorem~\ref{thm:herbrand-skolem} follows from Lemmata~\ref{ap1} and~\ref{ap2} in the appendix.

@@

\subsection{Reducing the encoding}
